Sparse Function-space Representation \\ of Neural Networks

### 2023-05-15 paper/main.tex
fixed typos

% Introduce the laplace GGN + shortcomings
%\textbf{Function-space methods}
One common approach for uncertainty with neural networks is a Laplace-GGN approximation \citep{daxberger2021laplace}, which takes a trained neural network and linearises it around the optimal weights. The Hessian can be efficiently approximated using the generalised Gauss--Newton approximation \cite{botev2017practical} but typically involves a cubic scaling in the number of parameters, in practice a further approximation such as the Kronecker factorisation \cite{martens2015optimizing,ritter2018kfac} is needed. The resulting linear model can be used to obtain uncertainty estimates and refine the neural network predictions \citep{immer2021scalable}, and is linear with respect to the weights. Therefore, as shown in \cite{immer2021scalable, khan2019approximate,maddox2021fast}, the linear model can be converted to a GP, which is then used to obtain the uncertainty estimates. 
%
However, the GP introduces a cubic scaling $\mathcal{O}(N^3)$ in the number of data points $N$. Previous approaches apply crude approximations to alleviate the computational costs, \eg, by considering only a subset of the training data \cite{immer2021scalable}. In the GP community, the scaling problem is not seen as an issue any more due to efficient and stochastic methods for sparse approximations (\eg, \cite{hensman2013gaussian,wang2019exact}). However, it is not entirely clear how to combine the linearization of the neural network with sparse methods \cite{titsias2009variational} without falling back to crude subset methods. We tackle this problem by framing the problem in the dual parameters \cite{csato2002sparse} of the GP that has previously been used for non-conjugate likelihood models in GPs \cite{adam2021dual}.

@@

\textbf{Linearization gives rise to a Gaussian process}
Our goal is to build a sparse function-space representation of a neural network.
Gaussian processes $f$ are an extension of the multivariate normal that characterizes a distribution over
functions in terms of the first two moments.
They are defined by a mean function $\mu(\cdot)$ and a covariance (aka kernel) function $\kappa(\cdot,\cdot)$.
Thus, we need these two functions based on the neural network to build the Gaussian process posterior.
Recent work \citep{khan2019approximate,maddox2021fast} has shown that linearizing approximations in the weight space
lead to function space equivalent approximations.
As Gaussian distributions remain tractable under linear transformations, a linear function in
terms of parameters can be converted from the weight space to the function space (see Ch.~2.1 in \cite{rasmussen2006gaussian}) as follows:\begin{equation} \label{eq:weight_func}
g_\weights(\mathbf{x}_{i}) = \phi^\top\!(\vx_i) \, \vw \implies \mu(\vx_i) = 0 \quad \kappa(\mathbf{x}_i, \mathbf{x}_i') = \frac{1}{\delta} \phi(\vx_i)^\top \phi(\vx_i).
\end{equation}
% The posterior structure directly relates to the optimization loss around the MAP weights $\vw^*$.
A common approach is to approximate the correlation structure of a distribution centred at the MAP estimate as done in the Laplace-GGN \citep{khan2019approximate, daxberger2021laplace, maddox2021fast}. The Laplace-GGN takes the MAP solution and approximates the Hessian of the loss function
% $\ell(f_\weights(\mathbf{x}_{i}), \mathbf{y}_i)$
with the Generalized Gaussian-Newton approximation (GGN).
Following the work of \citep{khan2019approximate}, we can view this approximation as building an approximate linear model of the neural network as $\hat{f}_\weights(\mathbf{x}_i) = \Jac{\weights_*}{\mathbf{x}_i} \, \weights$, where $\Jac{\weights}{\mathbf{x}_i} \coloneqq \nabla_\weights f(\mathbf{x}_i)^\top \in \R^{C \times P}$ is the Jacobian at the MAP.
Using the Hessian of the approximate model, we arrive at the Laplace-GGN posterior for $q(\vw)$.
\todo{do we need this sentence about the Hessian to get Laplace-GGN??}
Therefore, the GGN approximate linear model can be used to convert our weight space model to the function space by,
\begin{equation} 
\label{eq-laplace-approx-function-space}
% g(\vx) \sim \GP \left( \mu(\vx), \kappa(\vx, \vx') \right) \quad \text{with} \quad
  \mu(\vx_i) =  0 \quad \text{and} \quad
  \kappa(\mathbf{x}_i, \mathbf{x}_i')
  = \frac{1}{\delta} \Jac{\weights^*}{\mathbf{x}_i}\JacT{\weights^*}{\mathbf{x}_i'},
\end{equation}
where the kernel is the so-called Neural Tangent Kernel (NTK, \cite{jacot2018neural}). With this kernel function, we can then combine it with the dataset $\dataset$ to form the posterior $q(\vf)$. \citet{khan2019approximate} took a similar approach, but instead of fitting the GP posterior to the actual $\vy$, but a transformation of $\vy$ which they refer to as pseudo-data. Similarly, \citet{immerImprovingPredictionsBayesian2021} obtain the same covariance function but a different mean because they rely on a first order approximation of the neural network to form a Bayesian GLM model. Both approaches attempted to adjust the GP posterior mean function to ensure the predictions are from the neural network $f_{\vw^*}(\vx)$. In our method, we wish to make predictions from the derived GP and thus we avoid such adjustments. We now present the general formulation of the sparse functional representation of our trained neural network based on the non-sparse approximation.

>>>>

% Introduce the laplace GGN + shortcomings
%\textbf{Function-space methods}
One common approach for uncertainty with neural networks is a Laplace-GGN approximation \citep{daxberger2021laplace}, which takes a trained neural network and linearises it around the optimal weights. The Hessian can be efficiently approximated using the generalized Gauss--Newton approximation (GGN)~\cite{botev2017practical} but typically involves a cubic scaling in the number of parameters, in practice a further approximation such as the Kronecker factorisation \cite{martens2015optimizing,ritter2018kfac} is needed. The resulting linear model can be used to obtain uncertainty estimates and refine the neural network predictions \citep{immer2021scalable}, and is linear with respect to the weights. Therefore, as shown in \cite{immer2021scalable, khan2019approximate,maddox2021fast}, the linear model can be converted to a GP, which is then used to obtain the uncertainty estimates. 
%
However, the GP introduces a cubic scaling $\mathcal{O}(N^3)$ in the number of data points $N$. Previous approaches apply crude approximations to alleviate the computational costs, \eg, by considering only a subset of the training data \cite{immer2021scalable}. In the GP community, the scaling problem is not seen as an issue any more due to efficient and stochastic methods for sparse approximations (\eg, \cite{hensman2013gaussian,wang2019exact}). However, it is not entirely clear how to combine the linearization of the neural network with sparse methods \cite{titsias2009variational} without falling back to crude subset methods. We tackle this problem by framing the problem in the dual parameters \cite{csato2002sparse} of the GP that has previously been used for non-conjugate likelihood models in GPs \cite{adam2021dual}.

@@

\textbf{Linearization gives rise to a Gaussian process}
Our goal is to build a sparse function-space representation of a neural network.
Gaussian processes $f$ are an extension of the multivariate normal that characterizes a distribution over
functions in terms of the first two moments.
They are defined by a mean function $\mu(\cdot)$ and a covariance (aka kernel) function $\kappa(\cdot,\cdot)$.
Thus, we need these two functions based on the neural network to build the Gaussian process posterior.
Recent work \citep{khan2019approximate,maddox2021fast} has shown that linearizing approximations in the weight space
lead to function space equivalent approximations.
As Gaussian distributions remain tractable under linear transformations, a linear function in
terms of parameters can be converted from the weight space to the function space (see Ch.~2.1 in \cite{rasmussen2006gaussian}) as follows:\begin{equation} \label{eq:weight_func}
g_\weights(\mathbf{x}_{i}) = \phi^\top\!(\vx_i) \, \vw \implies \mu(\vx_i) = 0 \quad \kappa(\mathbf{x}_i, \mathbf{x}_i') = \frac{1}{\delta} \phi(\vx_i)^\top \phi(\vx_i).
\end{equation}
% The posterior structure directly relates to the optimization loss around the MAP weights $\vw^*$.
A common approach is to approximate the correlation structure of a distribution centred at the MAP estimate as done in the Laplace-GGN \citep{khan2019approximate, daxberger2021laplace, maddox2021fast}. The Laplace-GGN takes the MAP solution and approximates the Hessian of the loss function
% $\ell(f_\weights(\mathbf{x}_{i}), \mathbf{y}_i)$
with the GGN.
Following the work of \citep{khan2019approximate}, we can view this approximation as building an approximate linear model of the neural network as $\hat{f}_\weights(\mathbf{x}_i) = \Jac{\weights_*}{\mathbf{x}_i} \, \weights$, where $\Jac{\weights}{\mathbf{x}_i} \coloneqq \nabla_\weights f(\mathbf{x}_i)^\top \in \R^{C \times P}$ is the Jacobian at the MAP.
Using the Hessian of the approximate model, we arrive at the Laplace-GGN posterior for $q(\vw)$.
\todo{do we need this sentence about the Hessian to get Laplace-GGN??}
Therefore, the GGN approximate linear model can be used to convert our weight space model to the function space by,
\begin{equation} 
\label{eq-laplace-approx-function-space}
% g(\vx) \sim \GP \left( \mu(\vx), \kappa(\vx, \vx') \right) \quad \text{with} \quad
  \mu(\vx_i) =  0 \quad \text{and} \quad
  \kappa(\mathbf{x}_i, \mathbf{x}_i')
  = \frac{1}{\delta} \Jac{\weights^*}{\mathbf{x}_i}\JacT{\weights^*}{\mathbf{x}_i'},
\end{equation}
where the kernel is the so-called Neural Tangent Kernel (NTK, \cite{jacot2018neural}). With this kernel function, we can then combine it with the dataset $\dataset$ to form the posterior $q(\vf)$. \citet{khan2019approximate} took a similar approach, but instead of fitting the GP posterior to the actual $\vy$, but a transformation of $\vy$ which they refer to as pseudo-data. Similarly, \citet{immerImprovingPredictionsBayesian2021} obtain the same covariance function but a different mean because they rely on a first order approximation of the neural network to form a Bayesian GLM model. Both approaches attempted to adjust the GP posterior mean function to ensure the predictions are from the neural network $f_{\vw^*}(\vx)$. In our method, we wish to make predictions from the derived GP and thus we avoid such adjustments. We now present the general formulation of the sparse functional representation of our trained neural network based on the non-sparse approximation.
